### 2024-10-28 paper/3-motivation.tex
Add control flow of SpecTec prose

% an example of Wasm control flow
\textbf{Example 1}
\begin{verbatim}
  // infinite loop
  (loop (result i32) (i32.const 42) (br 0) (unreachable) end) (unreachable)
\end{verbatim}

@@

% exit label in official prose
There is also a special form of a behavior related to the block: an
\textit{exit label}.
\cref{fig:exit-label} is the \officialp of the exit label.
It is special because this behavior is not performed without an explicit
WebAssembly instruction.
Rather, the behavior is performed when \textbf{the end of a block is reached}
without control instructions or runtime error.
The behavior is that the label is popped from the stack, and the pc changes to
the point after the end of the block.

\begin{figure}[h!]
    \centerline{\includegraphics[width=15cm]{fig/exit}}
    \caption[Enter the caption title here]{exit label} \label{fig:exit-label}
\end{figure}

% an example of exit label
\textbf{Example 2}
\begin{verbatim}
  // exit label
  (loop (result i32) (i32.const 42) end) (f64.const 3.14)
\end{verbatim}

Example 2 is \texttt{loop} with a block that has only \texttt{i32.const}.
A label is pushed in the stack when a \texttt{loop} instruction is executed.
After \texttt{i32.const} is executed, the end of the block is reached.
As a result, exit label occurs so that the label is popped from the stack,
pc changes to the point after the end of block: \texttt{f64.const}.
Therefore, the code example above results in the value 42 and 3.14 pushed to
the stack.

Similar to the control flow using the block and the label, there are many other
control instructions and structures such as a \texttt{call} instruction that
calls a function with a frame and a \texttt{return} instruction that returns from
the function with the frame.

TODO: detailed explanation of \spectecp{} w/ fig

However, the \spectecp{} assumes that the \red{machine} takes WebAssembly
instructions one by one, and changes the internal state accordingly without pc.
This is because the \spectecp{} is generated automatically from the \red{dsl}
which uses rewrite rule to describe WebAssembly semantics.
As a result, rather than jump to the start of the loop, \spectecp{} stores the
\texttt{loop} instruction itself in the label, and executes it again if the
\texttt{br} instruction is executed.
Not only that, it doesn't have block structure.

don't remember end!

% official spec prose notation control structure
% seems to be pc-based
% AL from DSL ~ formal notation: no pc
% AL: instruction sequence input continously
% device exit semantics for AL
% Executable spec passes all wasm testcases
% weird, maybe not happened in realworld, but valid wasm code
% wrong!

>>>>

% an example of Wasm control flow in official prose
\textbf{Example 1}
\begin{verbatim}
  // infinite loop
  (loop (result i32) (i32.const 42) (br 0) (unreachable) end) (unreachable)
\end{verbatim}

@@

% exiting label in official prose
There is also a special form of a behavior related to the block: an
\textit{exiting label}.
\cref{fig:exiting-label} is the \officialp of the exiting label.
It is special because this behavior is not performed without an explicit
WebAssembly instruction.
Rather, the behavior is performed when \textbf{the end of a block is reached}
without control instructions or runtime error.
The behavior is that the label is popped from the stack, and the pc changes to
the point after the end of the block.

\begin{figure}[h!]
    \centerline{\includegraphics[width=15cm]{fig/exiting}}
    \caption[Enter the caption title here]{exiting label} \label{fig:exiting-label}
\end{figure}

% an example of exiting label in official prose
\textbf{Example 2}
\begin{verbatim}
  // exiting label
  (loop (result i32) (i32.const 42) end) (f64.const 3.14)
\end{verbatim}

Example 2 is \texttt{loop} with a block that has only \texttt{i32.const}.
A label is pushed in the stack when a \texttt{loop} instruction is executed.
After \texttt{i32.const} is executed, the end of the block is reached.
As a result, exiting label occurs so that the label is popped from the stack,
pc changes to the point after the end of block: \texttt{f64.const}.
Therefore, the code example above results in the value 42 and 3.14 pushed to
the stack.

% general control flow structure of official prose
Similar to the control flow using the block and the label, there are many other
control instructions and structures such as a \texttt{call} instruction that
calls a function with a frame and a \texttt{return} instruction that returns from
the function with the frame.
Not only that, there are the special behaviors such as a \textit{returning from
a function} that corresponds to the exiting label.

% control flow structure in SpecTec prose
However, the \spectecp{} assumes different a model to explain the control flow
of WebAssembly.
Rather than using the notion of pc, \spectecp{} assumes that WebAssembly
instructions are just given one by one.
This is because the \spectecp{} is generated automatically from the
\red{SpecTec dsl}.
\red{SpecTec dsl} uses rewrite rule to describe WebAssembly semantics, and it
can be seen as consuming a WebAssembly instruction and values in the stack,
changing the stack and the store, and inputing new WebAssembly instructions.

% an example of Wasm control flow in official prose
\begin{figure}[h!]
    \centerline{\includegraphics[width=15cm]{fig/spectec-loop}}
    \caption[Enter the caption title here]{SpecTec \texttt{loop}} \label{fig:spectec-loop}
\end{figure}
\begin{figure}[h!]
    \centerline{\includegraphics[width=15cm]{fig/spectec-br}}
    \caption[Enter the caption title here]{SpecTec \texttt{br}} \label{fig:spectec-br}
\end{figure}

Consider example 1 above.
\texttt{loop} and \texttt{unreachable} is the input instruction for the model.
\cref{fig:spectec-loop} is the \spectecp{} of the \texttt{loop} instruction.
Rather than storing the point to jump as a continuation in the label, it stores
the \texttt{loop} instruction itself and \textit{enters} the block.
Here, \textit{enter} means that it pushes the label in the stack and inputs the
instructions as inputs to execute.
As a result, after the \texttt{loop} is executed, \texttt{i32.const},
\texttt{br}, \texttt{unreachable}, and \texttt{unreachable} are remained as
inputs to the model.
When \texttt{i32.const} is excuted, it pushes 42 as \officialp{} does.
However, \texttt{br} behaves a bit differently.
\cref{fig:spectec-br} is the \spectecp{} of the \texttt{br} instruction.
It pops the label from the stack, removes the input instructions until the end
of the block, considers the loop instruction in the label as a new input
instruction, and executes it again.
Consequently, remaining inputs are \texttt{loop}, and \texttt{unreachable}.
Therefore, it explains the behavior in different point of view, but actually
describes the same thing.

% exiting label in SpecTec prose
However, this viewpoint is challenged by the exiting label.
As the block structure doesn't remain intact, it is hard to describe
\textbf{the end of the block}.
Furthermore, the exiting label is not performed by a specific WebAssembly
instruction, which makes it hard to model the behavior.
To handle this problem, an administrative instruction \texttt{end} is
introduced in \spectecp.
When entering a block, assume that there is an \texttt{end}

% official spec prose notation control structure
% seems to be pc-based
% AL from DSL ~ formal notation: no pc
% AL: instruction sequence input continously
% device exiting semantics for AL
% Executable spec passes all wasm testcases
% weird, maybe not happened in realworld, but valid wasm code
% wrong!
